Universal scaling for the ionization of biological molecules by highly charged ions

### 2020-05-27 zscaling/submitted/zscaling.tex
scaling manuscript with major revisions

The ionization of biological molecules by multicharged ions has gained increasing interest due to medical and environmental reasons~\cite{PhysMed}, from medical treatments~\cite{Mohamad2017,Solov2009,Denifl2011} to contaminant recognition in biological materials~\cite{water,ferrazdias}. Many semiempirical \citep{vera_prl2013} and theoretical efforts are currently being undertaken~\cite{MendezJPB20,Quinto20,ludde2019,ludde2018,ludde2016,Champion2012} to get reliable values for the ionization cross sections of these molecules.

@@

At intermediate impact energies, 
Janev and Presnyakov~\cite{janev1980} suggest $\sigma/Z$ versus $E/Z$ as the \textit{natural} reduced form of the ionization cross section $\sigma$ and the incident ion energy $E$. Much more recently, Montenegro and co--workers~\cite{dubois13,montenegro_pra13} proposed an alternative scaling by taking into account that  $\sigma$ is a function of $Z^2/E$. Their scaling, given by
\begin{equation}
    \sigma/Z^{\alpha}=f(E/Z^{2-\alpha}),
    \label{Montenegro}
\end{equation}
keeps the $Z^2/E$ relationship for any value of $\alpha$. 
In Ref.~\cite{dubois13}, the authors propose $\alpha=4/3$ for ionization of He and H$_2$ by different charged ions.

@@

The scaling was tested with the experimental data available by impact of different charged ions \cite{itoh2013,iriki2011,wolff2014,wang2016,tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,Rudd85,Luna_Li_water,DalCappello2009,Tribedi_O_water},
and also by electron impact at sufficiently high velocity \cite{rahman2016,bug2017,wolf2019,fuss2009}. As can be noted, most of the data in Fig.~\ref{fig:zreduced} confirms the present scaling, even for O$^{+8}$ in water~\cite{Tribedi_O_water}. However, the data of uracil by swift C, O and F ions in \cite{agnihotri2012,agnihotri2013} are too low compared with our CDW--SSM results, but also as compared with Itoh \textit{et al.} data \cite{itoh2013}, and with the CTMC calculations by Sarkadi \cite{sarkadi2016}.
The data for Li$^{+3}$ in water from Ref.~\cite{Luna_Li_water} also spreads out from the present theoretical curves for $E<600$ keV/amu.

>>>>

The ionization of biological molecules by multicharged ions has gained 
increasing interest due to medical and environmental 
reasons~\cite{PhysMed}, from medical treatments~\cite{Mohamad2017,
Solov2009,Denifl2011} to contaminant recognition in biological 
materials~\cite{water,ferrazdias}. Many semiempirical 
\citep{vera_prl2013} and theoretical efforts are currently being 
undertaken~\cite{MendezJPB20,Quinto20,ludde2019,ludde2018,ludde2016,
Champion2012} to get reliable values for the ionization cross sections 
of these molecules.

@@

At intermediate impact energies, Janev and Presnyakov~\cite{janev1980} 
suggest $\sigma/Z$ versus $E/Z$ as the \textit{natural} reduced form of 
the ionization cross section $\sigma$ and the incident ion energy $E$. 
Much more recently, Montenegro and co--workers~\cite{dubois13,
montenegro_pra13} proposed an alternative scaling by taking into 
account that  $\sigma$ is a function of $Z^2/E$. Their scaling, given 
by
\begin{equation}
    \sigma/Z^{\alpha}=f(E/Z^{2-\alpha}),
    \label{Montenegro}
\end{equation}
keeps the $Z^2/E$ relationship for any value of $\alpha$. 
In Ref.~\cite{dubois13}, the authors propose $\alpha=4/3$ for 
ionization of He and H$_2$ by different charged ions.

@@

The scaling was tested with the experimental data available by impact 
of different charged ions \cite{itoh2013,iriki2011,wolff2014,wang2016,
tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,Rudd85,
Luna_Li_water,DalCappello2009,Tribedi_O_water}, and also by electron 
impact at sufficiently high velocity \cite{rahman2016,bug2017,wolf2019,
fuss2009}. As can be noted, most of the data in Fig.~\ref{fig:zreduced} 
confirms the present scaling, even for O$^{+8}$ in 
water~\cite{Tribedi_O_water}. However, the data of uracil by swift C, 
O and F ions in \cite{agnihotri2012,agnihotri2013} are too low compared 
with our CDW--SSM results, but also as compared with Itoh 
\textit{et al.} data \cite{itoh2013}, and with the CTMC calculations by 
Sarkadi \cite{sarkadi2016}. The data for Li$^{+3}$ in water from 
Ref.~\cite{Luna_Li_water} also spreads out from the present theoretical 
curves for $E<600$ keV/amu.
